```{r}
library(ggplot2)
library(corrplot)
library(readr)
library(dplyr)
library(caret)
library(ranger)
library(Metrics)
```


```{r}
data <- read.csv("diabetes_binary_health_indicators_BRFSS2015.csv")
```

```{r}
head(data)
```
```{r}
str(data)
```

```{r}
summary(data)
```
```{r}
dim(data)
```

```{r}
colSums(is.na(data))
```
```{r}
par(mfrow=c(2,4))
hist(data$BMI)
hist(data$GenHlth)
hist(data$MentHlth)
hist(data$PhysHlth)
hist(data$Age)
hist(data$Income)
hist(data$Education)
```
```{r}
options(dplyr.summarise.inform = FALSE)
counts <- data %>% group_by(Diabetes_binary, HighBP) %>% summarise(count = n())
counts
```

```{r}
#remove duplicate and check correlation
data <- subset(data, !duplicated(data))

# Filter binary and continuous columns
binary_columns <- data %>%
  select_if(function(x) all(x %in% c(0, 1)))

continuous_columns <- data %>%
  select_if(function(x) is.numeric(x) && !all(x %in% c(0, 1)))

# Print the filtered columns
dependent = binary_columns$Diabetes_binary
binary_columns <- binary_columns[, -which(names(binary_columns) == "Diabetes_binary")]

ordinal_columns <- subset(continuous_columns, select = c(Education, Income, Age))
continuous_columns <- subset(continuous_columns, select = -c(Education, Income, Age))
```

```{r}
print("Binary - Binary:")
for (var in colnames(binary_columns)) {
  cross_table <- table(dependent, binary_columns[[var]])
  chi_square <- chisq.test(cross_table)$statistic
  n <- sum(cross_table)
  phi_coefficient <- sqrt(chi_square /n)
  cat("Column name:", var, "\n")
  cat("Chi-square:", chi_square, "\n")
  cat("Phi coefficient:", phi_coefficient, "\n\n")
}
```


```{r}
library(ltm)
print("Binary - Ordinal:")
corr_res <- ltm(binary_columns ~ continuous_columns)
res <- ltm.biserial(corr_res)

for (var in colnames(ordinal_columns)) {

  correlation_result <- ltm(continuous_columns[[var]], dependent$binary_columns.Diabetes_binary)

  rank_biserial <- ltm.biserial(correlation_result)
  
  print(paste("Column name:", var))
  print(paste("rank biserial:", rank_biserial))
}
```
```{r}
#redeclare data for usage
binary_columns <- data %>%
  select_if(function(x) all(x %in% c(0, 1)))

continuous_columns <- data %>%
  select_if(function(x) is.numeric(x) && !all(x %in% c(0, 1)))

dependent = data.frame(binary_columns$Diabetes_binary)
binary_columns <- binary_columns[, -which(names(binary_columns) == "Diabetes_binary")]

ordinal_columns <- subset(continuous_columns, select = c(Education, Income, Age))
continuous_columns <- subset(continuous_columns, select = -c(Education, Income, Age))


print("Binary - continuos:")
for (var in colnames(continuous_columns)) {
  print(paste("Column name:", var))
  print("point biserial:")
  print(cor.test(continuous_columns[[var]], dependent$binary_columns.Diabetes_binary))
}
```

```{r}
table <- table(data$Diabetes_binary,data$GenHlth)
chisq.test(table)
```
X-squared represents the overall discrepancy between both variables, but doesn't provide any information about both of the variables association.
df or The degrees of freedom (df) represent the number of values that are free to vary in the calculation of the chi-square statistic.
The p-value is a measure of the strength of evidence against the null hypothesis.

since the p-value is < 0.05 ,we can conclude that Diabetes_binary has a strong relationship with General health. 

```{r}
indices <- createDataPartition(data$Diabetes_binary, p = 0.7, list = FALSE)
train <- data[indices, ]
test <- data[-indices, ]

model <- ranger(train$Diabetes_binary ~ ., data = train, num.trees = 100)

rmse(
  test$Diabetes_binary, 
  predict(model, test)$predictions
)

summary(model)

```




