```{r}
library(ggplot2)
library(corrplot)
library(readr)
library(dplyr)
library(caret)
library(ranger)
library(Metrics)
```


```{r}
data <- read.csv("diabetes_binary_health_indicators_BRFSS2015.csv")
```

```{r}
head(data)
```
```{r}
str(data)
```
```{r}
summary(data)
```
```{r}
dim(data)
```

```{r}
colSums(is.na(data))
```
```{r}
par(mfrow=c(2,4))
hist(data$BMI)
hist(data$GenHlth)
hist(data$MentHlth)
hist(data$PhysHlth)
hist(data$Age)
hist(data$Income)
hist(data$Education)
```
```{r}
options(dplyr.summarise.inform = FALSE)
counts <- data %>% group_by(Diabetes_binary, HighBP) %>% summarise(count = n())
counts
```

```{r}
#remove duplicate and check correlation
data <- subset(data, !duplicated(data))
correlations <- cor(data)
target_correlations <- correlations["Diabetes_binary", ]
target_correlations
```

```{r}

table <- table(data$Diabetes_binary,data$GenHlth)
chisq.test(table)

```
X-squared represents the overall discrepancy between both variables, but doesn't provide any information about both of the variables association.
df or The degrees of freedom (df) represent the number of values that are free to vary in the calculation of the chi-square statistic.
The p-value is a measure of the strength of evidence against the null hypothesis.

since the p-value is < 0.05 ,we can conclude that Diabetes_binary has a strong relationship with General health. 

```{r}
indices <- createDataPartition(data$Diabetes_binary, p = 0.7, list = FALSE)
train <- data[indices, ]
test <- data[-indices, ]

model <- ranger(train$Diabetes_binary ~ ., data = train, num.trees = 100)

rmse(
  test$Diabetes_binary, 
  predict(model, test)$predictions
)

summary(model)

```


